VOI-aware Monte-Carlo Sampling in Trees

### 2011-12-09 paper/vago/paper.tex
Monte Carlo => Monte-Carlo

\title{VOI-aware Monte Carlo Sampling in Trees}
\author {David Tolpin, Solomon Eyal Shimony \\
Department of Computer Science, \\
Ben-Gurion University of the Negev, Beer Sheva, Israel \\
\{tolpin,shimony\}@cs.bgu.ac.il}

@@

Efficient algorithms for Multi-Armed Bandits based on
distribution-independent bounds, in particular UCB1, are introduced in
\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to Monte
Carlo Tree Search is described in \cite{Kocsis.uct}, and a successful
application of UCT to playing the Go game is discussed in
\cite{Gelly.mogo}.

@@

The principles of bounded rationality appear in
\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
description of rational metareasoning and case studies of applications
in several problem domains. One obstacle to straightforward
application of the principles of rational metareasoning to Monte Carlo
sampling is the metagreedy assumption, according to which samples must
be selected as though at most one sample can be taken before an action
is chosen. In Monte Carlo sampling, the value of information of any
single sample in a given search state is often zero, so a different
approximating assumption must be used instead.

@@

The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
in the regret compared to selecting an arm without pulling any arm at
all. The \textit{myopic} VOI estimate is of limited applicability to
Monte Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be non-positive, resulting in premature
termination of the search. However, $\Lambda_i$ can be estimated as the intrinsic 
value of perfect information $\Lambda_i^p$ about the mean reward of the $i$th arm. Two
cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean is pulled, and the 
true mean of the arm is lower than the sample mean of the second-best arm $\beta$;
\item another arm is pulled, and the true mean of the arm is higher
than the sample mean of the best arm $\alpha$.
\end{itemize}
$\Lambda_i^p$ can be bounded from above as the probability that a
different arm is selected, multiplied by the
maximum possible increase in the reward:
\begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
  bounded from above as
\begin{equation}
  \Lambda_i^p \le \left\{
  \begin{array}{l l}
    \overline X_\beta \Pr(\IE[X_i] \le \overline X_\beta) & \mbox{{\rm if} $i=\alpha$} \\
    (1-\overline X_\alpha)\Pr(\IE[X_i] \ge \overline X_\alpha) & \mbox{\rm otherwise}
  \end{array} \right.
\label{eqn:thm-vopi}
\end{equation}
\label{thm:vopi}
\end{thm}

@@

The sampling policies are first compared on random multi-armed bandit 
problem instances, where the Monte Carlo Tree Search algorithm
is reduced to the sample selection policy at the first step: UCT is 
actually UCB1. In the problem of pure exploration in multi-armed bandits
\cite{Bubeck.pure}, the smaller is the exploration factor, the higher
is the expected simple regret; indeed, UCT exhibits the lowest simple
regret in the experiments for the default exploration factor $Cp=\frac
{\sqrt 2} 2$ \cite{Kocsis.uct}.

@@

For the experiments, a modified version of Pachi \cite{Braudis.pachi},
a state of the art Go program, was used:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping condition for the VOI-aware policies was
  modified and based solely on the sample cost, specified as
  a constant parameter. 
\item The time-allocation mode based on the fixed number of samples
  was modified such that 
  \begin{enumerate}
    \item the same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item if samples were unused at the current step,
      they become available at the next step.
  \end{enumerate}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is still
a strong player. On the other hand, additional
features of more advanced engines would obstruct the Monte Carlo
sampling phenomena which are the subject of the experiment.

@@

This work suggested Monte Carlo sampling policies in which sample
selection and termination are based on upper bounds on the value of
information. A simplified model of accounting for future value of
information of a sample, based on early termination and sample
redistribution, was proposed for Monte Carlo sampling in
trees. Empirical evaluation showed that the policies outperform
heuristic algorithms for simple regret minimization in multi-armed
bandits, as well as for tree search.

Monte Carlo tree search still remains a largely unexplored field of
application of VOI-aware algorithms. More elaborated VOI estimates,
taking into consideration re-use of samples in future search states
should be considered. The policies introduced in the paper differ from
the UCT algorithm only at the first step, where the VOI-aware
decisions are made. Consistent application of principles of rational
metareasoning at all steps of a rollout may further improve the
sampling policies.

>>>>

\title{VOI-aware Monte-Carlo Sampling in Trees}
\author {David Tolpin, Solomon Eyal Shimony \\
Department of Computer Science, \\
Ben-Gurion University of the Negev, Beer Sheva, Israel \\
\{tolpin,shimony\}@cs.bgu.ac.il}

@@

Efficient algorithms for Multi-Armed Bandits based on
distribution-independent bounds, in particular UCB1, are introduced in
\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and a successful
application of UCT to playing the Go game is discussed in
\cite{Gelly.mogo}.

@@

The principles of bounded rationality appear in
\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
description of rational metareasoning and case studies of applications
in several problem domains. One obstacle to straightforward
application of the principles of rational metareasoning to Monte-Carlo
sampling is the metagreedy assumption, according to which samples must
be selected as though at most one sample can be taken before an action
is chosen. In Monte-Carlo sampling, the value of information of any
single sample in a given search state is often zero, so a different
approximating assumption must be used instead.

@@

The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
in the regret compared to selecting an arm without pulling any arm at
all. The \textit{myopic} VOI estimate is of limited applicability to
Monte-Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be non-positive, resulting in premature
termination of the search. However, $\Lambda_i$ can be estimated as the intrinsic 
value of perfect information $\Lambda_i^p$ about the mean reward of the $i$th arm. Two
cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean is pulled, and the 
true mean of the arm is lower than the sample mean of the second-best arm $\beta$;
\item another arm is pulled, and the true mean of the arm is higher
than the sample mean of the best arm $\alpha$.
\end{itemize}
$\Lambda_i^p$ can be bounded from above as the probability that a
different arm is selected, multiplied by the
maximum possible increase in the reward:
\begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
  bounded from above as
\begin{equation}
  \Lambda_i^p \le \left\{
  \begin{array}{l l}
    \overline X_\beta \Pr(\IE[X_i] \le \overline X_\beta) & \mbox{{\rm if} $i=\alpha$} \\
    (1-\overline X_\alpha)\Pr(\IE[X_i] \ge \overline X_\alpha) & \mbox{\rm otherwise}
  \end{array} \right.
\label{eqn:thm-vopi}
\end{equation}
\label{thm:vopi}
\end{thm}

@@

The sampling policies are first compared on random multi-armed bandit 
problem instances, where the Monte-Carlo Tree Search algorithm
is reduced to the sample selection policy at the first step: UCT is 
actually UCB1. In the problem of pure exploration in multi-armed bandits
\cite{Bubeck.pure}, the smaller is the exploration factor, the higher
is the expected simple regret; indeed, UCT exhibits the lowest simple
regret in the experiments for the default exploration factor $Cp=\frac
{\sqrt 2} 2$ \cite{Kocsis.uct}.

@@

For the experiments, a modified version of Pachi \cite{Braudis.pachi},
a state of the art Go program, was used:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping condition for the VOI-aware policies was
  modified and based solely on the sample cost, specified as
  a constant parameter. 
\item The time-allocation mode based on the fixed number of samples
  was modified such that 
  \begin{enumerate}
    \item the same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item if samples were unused at the current step,
      they become available at the next step.
  \end{enumerate}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is still
a strong player. On the other hand, additional
features of more advanced engines would obstruct the Monte-Carlo
sampling phenomena which are the subject of the experiment.

@@

This work suggested Monte-Carlo sampling policies in which sample
selection and termination are based on upper bounds on the value of
information. A simplified model of accounting for future value of
information of a sample, based on early termination and sample
redistribution, was proposed for Monte-Carlo sampling in
trees. Empirical evaluation showed that the policies outperform
heuristic algorithms for simple regret minimization in multi-armed
bandits, as well as for tree search.

Monte-Carlo tree search still remains a largely unexplored field of
application of VOI-aware algorithms. More elaborated VOI estimates,
taking into consideration re-use of samples in future search states
should be considered. The policies introduced in the paper differ from
the UCT algorithm only at the first step, where the VOI-aware
decisions are made. Consistent application of principles of rational
metareasoning at all steps of a rollout may further improve the
sampling policies.
